---
title: "Assignment3- Decision_tree validation model"
output: html_notebook
---
Reference practice 7



```{r}
# Load required packages
library(caret)
library(pROC)
library(ggplot2)
library(dplyr)
library(rpart)

```


```{r}
# Load data
data <- read.csv("a3_dataset_drop_encoded.csv")
```



```{r}
data <- subset(data, select = -c(1,2,4,6,9,10,11,12,13,14,15,16,17,18,19,23,30,31))
```



```{r}
str(data)
```



Decision Tree




```{r}
categorical_columns <- c("PotentialFraud", "Race")

# Convert these columns to factors
data[categorical_columns] <- lapply(data[categorical_columns], as.factor)

# Verify the structure of the data to ensure the conversion is correct
str(data)

```


```{r}
# Data Partitioning
# Create training and test sets
set.seed(1000)
data.class <- data
train_index <- sample(1:nrow(data.class), 0.7 * nrow(data.class))
train <- data.class[train_index, ]
test <- data.class[-train_index, ]

list( train = summary(train), test = summary(test) )
```

method = "class": This specifies that the model is a classification tree. 
If it were a regression tree, you would use method = "anova".

cp=0: This is the complexity parameter. It’s used to control the size of the decision tree and to prevent overfitting. A smaller cp will lead to a larger tree, which might fit the training data more closely, but may not generalize well to new data. In this case, setting cp=0 means the tree will keep splitting until all leaves are pure or until the splits do not improve the fit by a factor of cp.

default maxdepth = 30, minsplit = 20, cp = 0.01

```{r}
# Exploring Parameters


# Define the formula for the model
formula <- PotentialFraud ~ . 

# Build the decision tree with parameters
# Complexity Parameter, is used to control tree complexity, hence preventing overfitting. A smaaler cp will lead to a larger tree.
# minsplit control th minimum number of observations that must exist in a node in order for a split to be attemped. A larger minsplit will result in fewer splits, leading to a simpler.
# maxdepth control the maximum depth of any node of the final tree, with the root node counted as depth 0. Depth is length of the longest path from a tree node to the root. The smaller maxdepth will result i a simpler model.

tree_model <- rpart(formula, data = train, method = "class",
                    control = rpart.control(cp = 0.01, maxdepth = 10, minsplit = 20))

```

```{r}
# Predict on the test set
predictions <- predict(tree_model, test, type = "class")

```

```{r}
# Evaluate the model
confusionMatrix(predictions, test$PotentialFraud)
```

```{r}
# Visualize the tree structure
printcp(tree_model)
```


```{r}
plotcp(tree_model)
```


```{r}


tree_model <- rpart(formula, data = train, method = "class",
                    control = rpart.control(cp = 0.001, maxdepth = 10, minsplit = 10))

```


```{r}
plotcp(tree_model)
```


```{r}
# Predict on the test set
predictions <- predict(tree_model, test, type = "class")

```

```{r}
# Evaluate the model
confusionMatrix(predictions, test$PotentialFraud)
```
```{r}
# Exploring Parameters


tree_model <- rpart(formula, data = train, method = "class",
                    control = rpart.control(cp = 0.001, maxdepth = 10, minsplit = 5))

```



```{r}
# Predict on the test set
predictions <- predict(tree_model, test, type = "class")
# Evaluate the model
confusionMatrix(predictions, test$PotentialFraud)
```



```{r}
# Exploring Parameters


tree_model <- rpart(formula, data = train, method = "class",
                    control = rpart.control(cp = 0.001, maxdepth = 5, minsplit = 5))

```



```{r}
# Predict on the test set
predictions <- predict(tree_model, test, type = "class")
# Evaluate the model
confusionMatrix(predictions, test$PotentialFraud)
```

```{r}
plotcp(tree_model)
```




```{r}

# Calculate performance indicators
library(caret)
pred <- predict(tree_model, newdata = test, type = "class")
conf_mat <- confusionMatrix(pred, test$PotentialFraud)
print(conf_mat$overall) # accuracy, kappa, etc.
print(conf_mat$byClass) # sensitivity, specificity, etc.


```



```{r}
# Discuss the most predictive variables
print(varImp(tree_model))

```







```{r}
library(rattle)
library(tibble)
library(bitops)
# Plot the tree
fancyRpartPlot(tree_model, palettes = c("Greens", "Reds"), sub = "")
```



```{r}
# Build the decision tree with parameters
tree_model.full <- rpart(formula, data = train, method = "class",
                    control = rpart.control(cp = 0, maxdepth = 5, minsplit = 20))



fancyRpartPlot(tree_model.full, palettes = c("Greens", "Reds"), sub = "")
```



```{r}
# Prune decision tree
p.tree.prune <- prune(tree_model.full, cp = tree_model.full$cptable[which.min(tree_model.full$cptable[,"xerror"]),"CP"])
```
```{r}
plotcp(p.tree.prune)

```
```{r}
# Prune decision tree
min_cp <- tree_model.full$cptable[which.min(tree_model.full$cptable[,"xerror"]),"CP"]
p.tree.prune <- prune(tree_model.full, cp = min_cp)

# Print the cp value
print(paste("The cp value used for pruning is:", min_cp))

```


```{r}
fancyRpartPlot(p.tree.prune, palettes = c("Greens", "Reds"), sub = "")
```

 
```{r}
# Make a prediction
fraud.predict <- predict(p.tree.prune, test, type = "class")
```

```{r}
# Print confusion matrix
table(fraud.predict, test$PotentialFraud)
```

```{r}
confusion_table <-table(fraud.predict, test$PotentialFraud)
confusion_table
```


```{r}
correct <- confusion_table |> diag() |>sum()
correct
```

```{r}
error <- confusion_table |>sum() - correct
error
```

```{r}
accuracy <- correct /(correct + error )
accuracy
```

```{r}
# function for accuracy
accuracy <- function(truth, prediction){
  tbl <- table(truth, prediction)
  sum(diag(tbl))/sum(tbl)
}
accuracy(test$PotentialFraud, fraud.predict)
```

Get a confusio table with more statistics (using caret)
```{r}
confusionMatrix(data=fraud.predict, reference = test |>pull(PotentialFraud))
```


Task 3. The ROC curve

+  pre-processing:

evaluating a classification model, particularly a decision tree model (p.tree.prune)

It's a binary classifier since you're using the second column of predicted probabilities.

type = "prob": This argument tells the predict() function to output probabilities rather than class predictions. For classification models, this allows you to get the probability of each class for each observation.

[ ,2]: This part of the code indicates that you're extracting the probabilities of the second class.


```{r}
# Calculating the probabilities of the positive class for the observations in the test dataset according to the decision tree model 
# 0: negative class
# 1: positive class

prob.fraud = predict(p.tree.prune, newdata = test, type = "prob")[,"1"] 

```

using roc() function from package 'pROC' to compute the Receiver Operating Characteristic (ROC) curve.

```{r}
res.roc <- roc(test$PotentialFraud, prob.fraud)
```

plot AUC: Area Under the Curve.

```{r}

plot.roc(res.roc, print.auc = TRUE)
```

```{r}
# Adding the best threshold value
plot.roc(res.roc, print.auc = TRUE, print.thres = "best")
```

Model Evaluation
ROC: Receiver Operating Characteristic
ROC curve plots False positive rate (x-axis)  against true positive rate (y-axis)

```{r}
# Extract thresholds, sentitivities, specificities
roc.data <- tibble(
  thresholds = res.roc$thresholds,
  sensitivity = res.roc$sensitivities,
  specificity = res.roc$specificities
)
roc.data

```


```{r}
# Get the probality threshold for specificity = 0.6
roc.data %>% filter(specificity >= 0.6)
```


Task 4. Cross validation


trainControl() is used to choose how testing is performed.

```{r}
# Loading required packages
library(caret)
library(rpart.plot)
```

K-fold cross-validation
defining training control as cross-validation and value of K equal to 10

```{r}
# Set up caret to perform 10-fold cross validation
cv.control <- trainControl(method = "cv",
number = 10)

```


rpart(formula, data, weights, subset, na.action = na.rpart, method,
      model = FALSE, x = FALSE, y = TRUE, parms, control, cost, ...)
      
      
The package caret combines training and validation for hyperparameter tuning into a single function called train(). It internally splits the data into training and validation sets and thus will provide you with error estimates for different hyperparameter settings.

```{r}
# Use caret to train the rpart decision tree using 10-fold cross
# validation and use 15 values for tuning the cp parameter for rpart.
# This code returns the best model.
rpart.cv <- train(PotentialFraud ~ .,
data = train,
method = "rpart",
trControl = cv.control,
tuneLength = 15)
```

```{r}
rpart.cv
```


Combine 2 command above in 1:
For rpart, train tries to tune the cp parameter (tree complexity) using accuracy to chose the best model.
Parameters used for tuning (in this case cp) need to be set using a data.frame in the argument tuneGrid! Setting it in control will be ignored.

```{r}
rpart.cv1 <- train(PotentialFraud ~  mean_InscClaimAmtReimbursed, mean_TotalReimbursementAmt, mean_TotalDeductibleAmt, State, County, AdmissionPeriod,
                    data = train,
                    method = "rpart",
                    
                    trControl = trainControl(method = "cv", number = 10),
                    tuneLength = 15)
rpart.cv1
```


```{r}
# Plot model selection
plot((rpart.cv1))
plot(rpart.cv)
```

A model using the best tuning parameters and using all the data supplied to train() is available as fit$finalModel.
extra = 2 : box wider *2

```{r}
rpart.plot(rpart.cv1$finalModel, extra = 2,
  box.palette = list("Gy", "Gn", "Bu", "Bn", "Or", "Rd", "Pu"))
```
The variable importance without competing splits.
```{r}
imp <- varImp(rpart.cv1, compete = FALSE)
imp
```
```{r}
ggplot(imp)
```







